hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2025-02-28 inst/manuscript/paper.Rmd
edits to checking data

To help users identify potential issues, **hatchR** includes basic data checking functions that highlight outliers or missing values both visually and programattically.

We demonstrate the utility of the summary and check functions `summarize_temp()`, `plot_check_temp()`, and `check_continuous()` using a simulated year-long data set (`year_sim`, the dimensions of the data and first 6 rows are shown) with temperature readings every thirty minutes.

@@

First, we recommend checking imported data for any outliers or strange inputs using `plot_check_temp()` (Figure 2).

@@

There are no obvious outliers but since each day has 48 records, we need to summarize it to daily mean temperature with `summarize_temp()` and then check for missing days with `check_continuous()`. We also recommend using `plot_check_temp()` again on the summarized data (though leave out the resulting plot for space efficiency in this manuscript).

>>>>

To help users identify potential issues, **hatchR** includes basic data checking functions that highlight outliers or missing values both visually and programatically. These checks ensure data integroty before model application.

We demonstrate the utility three functions: `summarize_temp()`, `plot_check_temp()`, and `check_continuous()`--using a simulated year-long data set (`year_sim`). This dataset contains temperature readings taken every thirty minutes, and its structure (dimentions and first six rows) is shown below.

@@

First, we recommend using `plot_check_temp()` to visually inspect imported data for outliers or unusual values (Figure 2).

@@

In this case, no obvious outliers are present, but since each day contains 48 records, the data must be summarized to daily mean temperature using `summarize_temp()`. After summarization, `check_continuous()` should be used to identify any missing days. We also suggest running `plot_check_temp()` again on the summarized data to verify its integrity, though we omit the resulting plot here for space efficiency.
